---
title: 'Prediction assignment: Write up'
author: "Betsega"
date: "November 2, 2019"
output:
  html_document: default
  pdf_document: default
---
## Summary
This progject aims at predicting the manner in which participants did a weight lifting excercise. A machine learing approach is applied
to build different models and model selction was done to get the best predictive approach

```{r setup, include=TRUE}

knitr::opts_chunk$set(echo = TRUE)

```



## Loading required libraries

```{r, echo=TRUE}

knitr::opts_chunk$set(echo = FALSE)

library(lattice)

library(ggplot2)

library(caret)

library(rpart) 

library(rpart.plot)

library(rattle)

library(randomForest)

library(RColorBrewer)

```
##loading data
```{r, echo=TRUE}


url_train <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"

url_quiz  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

data_train <- read.csv(url(url_train), strip.white = TRUE, na.strings = c("NA",""))

data_quiz  <- read.csv(url(url_quiz),  strip.white = TRUE, na.strings = c("NA",""))


dim(data_train)

dim(data_quiz)


```
### partition data into training and testing sets
```{r, echo=TRUE}
in_train  <- createDataPartition(data_train$classe, p=0.75, list=FALSE)
train_set <- data_train[ in_train, ]
test_set  <- data_train[-in_train, ]

dim(train_set)

dim(test_set)

```

## Remove the variables zeros

```{r, echo=TRUE}
nzv_var <- nearZeroVar(train_set)

train_set <- train_set[ , -nzv_var]

test_set  <- test_set [ , -nzv_var]


```
## Remove variables that are mostly NA. A threshlod of 95 % is selected.
```{r, echo=TRUE}
na_var <- sapply(train_set, function(x) mean(is.na(x))) > 0.95

train_set <- train_set[ , na_var == FALSE]

test_set  <- test_set [ , na_var == FALSE]

```

## Remove columns 1 to 5 since they are only identification varaibles


```{r, echo=TRUE}

train_set <- train_set[ , -(1:5)]

test_set  <- test_set [ , -(1:5)]


```
#Building Predictive Models

## 1 A decision tree model


```{r, echo=TRUE}

set.seed(2000)

fit_decision_tree <- rpart(classe ~ ., data = train_set, method="class")

fancyRpartPlot(fit_decision_tree)

```
### apply the decsion tree model to predict with the test data set

```{r, echo=TRUE}
predict_decision_tree <- predict(fit_decision_tree, newdata = test_set, type="class")

conf_matrix_decision_tree <- confusionMatrix(predict_decision_tree, test_set$classe)

conf_matrix_decision_tree

```

##2. Random Forest Model
```{r, echo=TRUE}
set.seed(2000)
ctrl_RF <- trainControl(method = "repeatedcv", number = 5, repeats = 2)
fit_RF  <- train(classe ~ ., data = train_set, method = "rf",
                 trControl = ctrl_RF, verbose = FALSE)
                 
fit_RF$finalModel

```
### Apply the randomForest model to predct with the test data set
```{r, echo=TRUE}
predict_RF <- predict(fit_RF, newdata = test_set)
conf_matrix_RF <- confusionMatrix(predict_RF, test_set$classe)
conf_matrix_RF

```
## Selecting the best model

 The predictive accuracy of the ttwo models evaluated is as follows:
  
   Decision Tree Model: 74.90 %

   Random Forest Model: 99.80 %

   Random Forest model is best 

## applying the selected model to predict the 20 data points (data_quize set)
```{r, echo=TRUE}
predict_quiz <- predict(fit_RF, newdata = data_quiz)
predict_quiz

```

